\documentclass[12pt]{letter}

\usepackage[letterpaper]{geometry}

\begin{document}

\begin{letter}{Editor, \textit{Journal of the Electrochemical Society
}}

  \opening{Dear Editor,}

  We are pleased to submit this manuscript for consideration in the \textit{Journal of the Electrochemical Society}. We confirm that this work has not been submitted to any other journal.
  
  In this work, we investigate ``knees'' in lithium-ion battery aging trajectories. We review prior work and identify six knee ``pathways'', including lithium plating, electrode saturation, resistance growth, additive depletion, percolation-limited connectivity, and mechanical deformation. We discuss which of these mechanisms are identifiable from experimental data. We also identify key design and usage sensitivities for knees.
  
  Finally, we discuss challenges and opportunities for knee modeling and prediction in both the laboratory and the field. Our findings illustrate the complexity and subtlety of lithium-ion battery degradation and can aid both academic and industrial efforts to predict and mitigate knees, and hence improve battery lifetime.
  
  Thank you for considering our submission.
  
  \closing{Regards,
  
  Peter Attia}

  \ps On behalf of Alexander Bills, Ferran Brosa Planella, Philipp Dechent, Gonçalo dos Reis, Matthieu Dubarry, Paul Gasper, Richard Gilchrist, Samuel Greenbank, David Howey, Ouyang Liu, Edwin Khoo, Yuliya Preger, Abhishek Soni, Shashank Sripad, Anna G. Stefanopoulou, and Valentin Sulzer

\end{letter}

\end{document}